---
title: "Ten examples using dplyr::select"
author: "Allison Horst"
date: "5/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use **dplyr::select** to select, de-select, or re-order columns in a data frame. 

Documentation: [dplyr::select](https://dplyr.tidyverse.org/reference/select.html)
Select helpers: [click here](https://www.rdocumentation.org/packages/tidyselect/versions/0.2.5/topics/select_helpers)

Remember: [dplyr::slice()](https://dplyr.tidyverse.org/reference/slice.html) for rows selection/exclusion

Load the tidyverse:
```{r, warning = FALSE}
suppressMessages(library(tidyverse)) # load tidyverse & suppress messages
```

Make some mock data:
```{r}

df <- tribble(
  ~red, ~orange, ~yellow, ~green, ~blue,
  1, 2, 3, "yes", "bird",
  0, 5, 10, "no", "dog",
  NA, 20, NA, "yes", "cat"
)

```


### Example 1: Select one column

*Create new data frame df_1 that only contains column 'yellow'.*

```{r}

df_1 <- df %>% 
  select(yellow) # alternatively, select(df, yellow)

```

### Example 2: Select multiple sequential columns:

*Select columns orange through green:*

```{r}

df_2 <- df %>% 
  select(orange:green) # alternatively, select(df, orange:green)

```

### Example 3: Select multiple non-sequential columns:

*Select columns red and yellow:*

```{r}

df_3 <- df %>% 
  select(red, yellow) # alternatively, select(df, red, yellow)

```

### Example 4: Select a combination of sequential and non-sequential columns:

*Select columns orange and green through blue:*

```{r}

df_4 <- df %>% 
  select(orange, green:blue) # alternatively, select(df, orange, green:blue)

```

### Example 5: Remove column with a minus sign in select:

*Keep everything _except_ orange:*

```{r}

df_5 <- df %>% 
  select(-orange) # alternatively, select(df, -orange)

```

### Example 6: Select any column that contains a designated string in the name:

*Keep all columns where the string "re" shows up in the name (here, in "red" and "green"):*

```{r}

df_6 <- df %>% 
  select(contains("re"))

```

### Example 7: Keep columns that start with or end with a designated string:

*Keep only columns that end in "e" using ends_with (alternative is starts_with):*

```{r}

df_7 <- df %>% 
  select(ends_with("e"))

```


### Example 8: Keep selected and reorder

*Keep only columns green, blue, and orange in that order:*

```{r}

df_8 <- df %>% 
  select(green, blue, orange)

```


### Example 9. Use it in combination with other functions

*Only keep observations (rows) where 'green' matches "yes", then only keep columns green and red, then rename green to 'lime'.*

```{r}

df_9 <- df %>% 
  filter(green == "yes") %>% 
  select(green, red) %>% 
  rename(lime = green)

```

### Example 10. Another combination example

*Deselect columns if their name contains an "o", filter to only keep observations where red is not NA.*

```{r}

df_10 <- df %>% 
  select(-contains("o")) %>% 
  filter(!is.na(red))

```

### Future things to try: select_if, select_at, rename_if, rename_at, rename_all
