---
title: "Ten examples using dplyr::mutate + dplyr::case_when"
author: "Allison Horst"
date: "5/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- [dplyr::mutate](https://dplyr.tidyverse.org/reference/mutate.html): create variables (columns) 

- [dplyr::case_when](https://dplyr.tidyverse.org/reference/case_when.html): intuitive if/if-else (for example: every time the word "wow" shows up in this column, I want the word "positive" to show up in the new column I create with mutate)

Using dplyr::mutate and dplyr::case_when in combination I think of as "I want to add a new column, where the entries in that new column are conditional based on what exists (value or character) in an existing column."

Load the tidyverse:
```{r, warning = FALSE}
suppressMessages(library(tidyverse)) # load tidyverse & suppress messages
```

Make some mock data: 
```{r}
df <- tribble(
  ~col_a, ~col_b, ~col_c,
  "wow", 5, "dog",
  "nope", 2, "cat",
  "wow", 4, "dog",
  "ok", 3, "dog",
  "nope", 1, "dog",
  "wow", NA, "cat",
  "ok", 2, "cat"
)

```

## Ten examples of how to use dplyr::mutate + dplyr::case_when together in slightly different ways: 

### Example 1. Adding a new character column with entries based on an existing character column

*I want to add a new column (col_d) based on responses in existing col_a. If col_a is "wow" then col_d should be "positive"; if col_a is "ok", then col_d should be "neutral"; if col_a is "nope" then col_d should be "negative."*

```{r}

df_1 <- df %>% # Make a new data frame 'df_1', using data from 'df' and then...
  mutate( # Adding a new column called...
    col_d = case_when( # 'col_d', which contains conditional entries where...
      col_a == "wow" ~ "positive", # when col_a is "wow", col_d should be "positive"...
      col_a == "ok" ~ "neutral", # when col_a is "ok", col_d should be "neutral"...
      col_a == "nope" ~ "negative" # and when col_a is "nope", col_d should be "negative"
    )
  )

```

### Example 2. Adding a new character column with entries based on existing numeric column

*I want to add a new column 'col_d' to df, where if col_b is greater than 3, then col_d is "high"; if col_b is 3, then col_d is "middle"; and if col_b is less than 3, then col_d is "low".*

```{r}

df_2 <- df %>% # Take original data frame df, and then...
  mutate( # add a new column called....
    col_d = case_when( # col_d, which contains conditional entries where...
      col_b > 3 ~ "high", # if col_b > 3, col_d is "high"...
      col_b == 3 ~ "middle", # if col_b = 3, col_d is "middle" (notice double == here)...
      col_b < 3 ~ "low" # if col_b < 3, col_d is "low
    )
  )

```

### Example 3. Adding a numeric column based on info in an existing character column

*I want to make a new column 'col_d' where: if col_c is "cat", then col_d should be 1; if col_c is "dog", then col_d should be 2.*

```{r}

df_3 <- df %>% # Create a new data frame 'df_3', start with 'df' and then...
  mutate( # add a column...
    col_d = case_when( # called 'col_d' that contains conditional entries where...
      col_c == "cat" ~ 1, # if col_c is "cat", then col_d should be 1...
      col_c == "dog" ~ 2 # and if col_c is "dog", then col_d should be 2
    )
  )

```

### Example 4. Adding a numeric column that is *calculated* from an existing column(s)

*I want to create a new column col_d, which: if col_a is "wow", then col_d will be 10 x col_b; if col_a is "nope" or "ok", then col_d will be 100 x col_b.*

```{r}

df_4 <- df %>% # Create a new data frame 'df_4', starting with 'df' and then...
  mutate( # add a column...
    col_d = case_when( # called 'col_d' that contains conditional entries where...
      col_a == "wow" ~ col_b*10, # if col_a is "wow", then col_d should be 10*col_b...
      col_a == "ok" | col_a == "nope" ~ col_b*100 # if col_a is "ok" or "nope", then col_d = 100*col_b
    )
  )

```

### Example 5. Add a new column that contains "this" if one condition is met, or "that" in any other case

*If col_a is "ok", then col_d should contain "so-so"; if col_a is anything ELSE, then col_d should contain "not ok".*

```{r}

# Note: adding 'TRUE ~ ' at the end of your set conditions seems like (?) saying "and for any other scenarios that I haven't specified, set those to [this thing]."

df_5 <- df %>% # Create a new data frame 'df_5', starting with 'df' and then...
  mutate( # add a column...
    col_d = case_when( # called 'col_d' that contains conditional entries where...
      col_a == "ok" ~ "so-so", # if col_a is "ok", then col_d should be "so-so"...
      TRUE ~ "not ok" # if col_a is "wow" or "nope", then col_d is "not ok"
    )
  )

```

### Example 6. Assign a value / character to existing 'NA's

*I want to create a new column that is "present" if col_b is not NA, and "absent" if col_b is NA.*

```{r}

df_6 <- df %>% # Create a new data frame 'df_6', starting with 'df' and then...
  mutate( # add a column...
    col_d = case_when( # called 'col_d' that contains conditional entries where...
      !is.na(col_b) ~ "present", # if col_b is NOT NA, return "present" in col_d
      is.na(col_b) ~ "absent" # if col_b IS NA, return "absent" in col_d
    )
  )

```

### Example 7. New column entry based on conditions for multiple variables

*If col_a is "wow" AND col_c is "dog", then col_d should be "yes." Otherwise, col_d should be "no".*

```{r}

df_7 <- df %>% 
  mutate( 
    col_d = case_when(
      col_a == "wow" & col_c == "dog" ~ "yes",
      TRUE ~ "no"
    )
  )

```

### Example 8. Combining with other wranglings

*First I want to only keep observations where col_c is "dog," then select columns col_a through col_c, then add a column col_d that is twice col_b if col_a is "wow", otherwise 0.*

```{r}

df_8 <- df %>% 
  filter(col_c == "dog") %>% 
  select(col_a:col_c) %>% 
  mutate(
    col_d = case_when(
      col_a == "wow" ~ 2*col_b,
      TRUE ~ 0
    )
    )

```

### Example 9. Setting levels with case_when: operates sequentially, so you only need to give one endpoint for each bin.

```{r}

```

