---
title: "Ten examples of getting started with functions"
author: "Allison Horst"
date: "5/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Step 0: Read chapter on functions in R4DS

[https://r4ds.had.co.nz/functions.html](R4DS Ch. 19 - Functions)

Load the tidyverse:
```{r, warning = FALSE, message = "hide"}
suppressMessages(library(tidyverse))
```

### Example 1. The simplest little function with a single input

Create a function 'function_one' where you input an argument (x), and it calculates x^2 and returns the value.

```{r}

function_one <- function(x){ 
  squared <- x^2
  return(squared)
}

function_one(3) # Note that this returns 9
function_one(212) # And this returns 212^2 = 44944

```

### Example 2. A function with multiple inputs

```{r}

function_two <- function(a,b){
  burrito <- a + b
  return(burrito)
}

function_two(2,6)

```

### Example 3. A function with named variables

- Add as many arguments as you want in the function: function(arg_1, arg_2, arg_3)

- Remember to use thoughtful/logical/nice names (possibly descriptive, concise, snake_case)

- You can create variables outside the function that are called (good for running through vectors/matrices!)

```{r}

apple <- 10
banana <- 6

function_three <- function(apple, banana, orange){
  fruit <- apple/orange + banana
  return(fruit)
}

function_three(apple, banana, 2) # Notice that I give the two assigned variable names here, but then give it a value for 'orange' because that wasn't specified previously

# You can also specify values later on: 

function_three(5, banana, 2)

```

### Example 4. A function with vector input(s)

Sometimes you want to pass a bunch of values through a function and see what the outcomes are. That's fine! Make some vectors (matrices later...) and plug in those variables:

```{r}

dog <- c(1,2,3,20) # A vector called 'dog'
cat <- c(6,7,8,9) # A vector called 'cat'

function_four <- function(dog,cat){
  daily_pets <- dog*2 + cat/3
  return(daily_pets)
}

```

### Example 5. A function with inputs from a data frame / tibble

You can even get vectors from columns in a data frame! Here, I'll make some mock data, then feed the columns into a function:

```{r}
# Make a little tibble:
fun_df <- tribble(
  ~e,~f,~g,
  0,1,2,
  3,4,5,
  6,7,8
)

# Write a function:

function_five <- function(arg_1, arg_2){
  output <- arg_1 + arg_2
  return(output)
}

# And this calculates the sum of values in columns e and f, for each row:
function_five(arg_1 = fun_df$e, arg_2 = fun_df$f) # Returns: 1  7  13

# Note: yes, in this case it's way easier to just use dplyr::mutate()

```

### Example 6. A function with conditions (using switch())

Sometimes you'll want a function to do different operations but use a generally similar function. You can use the switch() function to specify different selections (similar to using an if-else statement).

In this example: I want to write a function that either calculates the ratio as a decimal or percentage for two input values.

*Note: this doesn't seem quite right...*

```{r}

function_six <- function(val_1,val_2,to_do){
  switch(to_do,
         calc_decimal = val_1/val_2,
         calc_percent = 100*val_1/val_2
  )
}

function_six(10,30,2)

```

