---
title: "Ten examples with NAs"
author: "Allison Horst"
date: "5/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note: for examples on assessing / visualizing missingness, see [naniar package](https://cran.r-project.org/web/packages/naniar/vignettes/getting-started-w-naniar.html). 

Load the tidyverse and janitor:
```{r, warning = FALSE, message = "hide"}
suppressMessages(library(tidyverse))
suppressMessages(library(janitor))
```

Make some mock data:
```{r}

df <- tribble(
  ~red, ~orange, ~yellow, ~green, ~blue, ~purple,
  NA, 2, 3, NA, "bird", NA,
  0, 5, 100, "no", "dog", NA, 
  NA, 20, NA, "yes", "cat", NA,
  100, 200, 300, "no", "bird", 4
)

```


### Example 1. Filter out observations (rows) with NA values for a selected variable

*Only keep rows where yellow is NOT 'NA':*

```{r}

df_1 <- df %>% 
  dplyr::filter(!is.na(yellow)) # Note exclamation point for "NOT"

```

### Example 2. Remove rows that contains _any_ NAs across variables

*Use tidyr::drop_na() to return rows with no NA values (only row 4 here):*

See also: na.omit()

See also: complete.cases

See also: ggplot2::remove_missing

```{r}

df_2 <- df %>% 
  tidyr::drop_na()

```

### Example 3. Remove columns/rows that **only** contains NA

*Use janitor::remove_empty() to get rid of columns/rows full of NAs:*
(yeah in this case there isn't one, so all columns are retained...)

```{r}

df_3 <- df %>% 
  janitor::remove_empty("cols") # Can switch to "rows" to remove rows with only NAs!

```

### Example 4. Set NA values to something else, by column

*If red is NA, set value to 500; if purple is NA, set to "yay" using tidyr::replace_na:*

```{r}

df_4 <- df %>% 
  tidyr::replace_na(list(red = 500, purple = "yay"))

```

### Example 5. Replace *all* NAs in a data frame with something

*Anywhere NA shows up, replace with "not recorded":*

```{r}

df_5 <- df %>% 
  replace(., is.na(.), "unrecorded") # Use '.' for across all columns...

```

### Example 6. Within columns, replace value/character with NA

*If column 'blue' is "bird", make that cell NA using dplyr::na_if:*

See also: naniar::replace_with_na()

Note: here, this column overwrites original 'blue' column; consider makign the column within mutate() a different name to add another column (vs. replace existing)

```{r}

df_6 <- df %>% 
  dplyr::mutate(
    blue = dplyr::na_if(blue, "bird")
  )

```

### Example 7. Across all columns, replace value / character with NAs

*Anywhere the value 100 shows up in the data frame, replace with NA:*

```{r}

df_7 <- df %>% 
  dplyr::na_if(., 100) # In my head, '.' indicates across all?

```

### Example 8. Remove any column with **any** NA values (not sure why you'd want to do this, but anyway...)

*Remove any columns containing at least one NA value using a combination of select_if, any, and is.na:*

```{r}

df_8 <- df %>% 
  dplyr::select_if(~ !any(is.na(.))) # Looks for NAs across all columns; says only select (keep) columns if there are NOT (!) any NAs (not even sure if I'm interpreting that weird syntax right but whatever)

```

### Example 9. Omit NAs from a vector for a calculation (or sometimes a plot, or whatever...)

```{r}

# Uh-oh version:
yellow_mean_na <- mean(df$yellow) # Returns NA because there is an NA in that column

# The thing you probably want: 
yellow_mean <- mean(df$yellow, na.rm = TRUE) # Returns 6.5 (omits NA in calculation)

# Similarly: 

yellow_mean_b <- df %>% 
  pull(yellow) %>% # Pulls out just the vector data (values/entries in that column)
  mean(na.rm = TRUE)  

```


### Example 10. NAs in ggplot2 graphs?

```{r}

ggplot(df, aes(x = orange, y = yellow)) +
  geom_point()

# Handles it: note warning "Removed 1 rows containing missing values (geom_point)."

```

