---
title: "Ten examples: readxl"
author: "Allison Horst"
date: "8/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The {readxl} package provides a straightforward way to import data from Excel (.xls or .xlxs) from one or multiple worksheets. Here are some places to start learning more about readxl: 

- https://readxl.tidyverse.org/

- https://rpubs.com/tf_peterson/readxl_import

```{r include=FALSE}
library(tidyverse)
library(readxl)
```

Here, I'll use the 'coral.xlsx' file (see information here: https://catalog.data.gov/dataset/effects-of-thermal-stress-and-nitrate-enrichment-on-the-larval-performance-of-two-caribbean-ree/resource/32942be2-4bfe-4e4a-9648-097525cee05e)

1. Reading in the first worksheet:
```{r}

coral_1 <- read_excel("coral.xlsx") # Note that this reads in the first worksheet (not subsequent). Yay! 

```

2. Reading in only some rows or columns from the first worksheet:
```{r}

# Read in rows 1 - 8 (note: this INCLUDES the header)
coral_2a <- read_excel("coral.xlsx", range = cell_rows(1:8))

# Read in columns 2 and 3
coral_2b <- read_excel("coral.xlsx", range = cell_cols(c(2,3)))

# Read in range A10:C15
coral_2c <- read_excel("coral.xlsx", range = "A1:C10")

```

3. To read in a single worksheet that is NOT the first one (by worksheet name or position)
```{r}

coral_3 <- read_excel("coral.xlsx", sheet = "Pastreoides respiration")

```

4. To read in a single worksheet by name, choose a range, and set column names: 
```{r}

coral_4a <- read_excel("coral.xlsx", 
                      sheet = "Pastreoides respiration", 
                      range = "A4:C10", 
                      col_names = c("nutrient_level","temp","ox_cons")
                      )

# Alternatively, reference by worksheet position (but this is risky, since they can be moved around...)

coral_4b <- read_excel("coral.xlsx",
                       sheet = 2,
                       range = "A4:C10",
                       col_names = c("nutrients","temp","ox"))

```

5. To read in multiple worksheets: 
```{r}

path <- "coral.xlsx"

coral_5 <- path %>% 
  excel_sheets() %>% 
  set_names() %>% 
  map_df(read_excel, 
      path = path)

# Question: This binds data from all 3 worksheets together in sequential rows...what if we want to have them be separate data frames? Any way to load as separate dfs? 

```

6. Reassigning a character (or whatever) to NA when read into R: 
```{r}

# Anywhere "Low" appears, replace in R with 'NA'
coral_6 <- read_excel("coral.xlsx", na = "Low") 

```

7. Skipping the first *n* rows (auto-excludes leading empty rows). This is a good way in case there is additional (unnecessary descriptive) information in the top *n* rows.
```{r}

coral_7 <- read_excel("coral.xlsx", skip = 3, col_names = c("one","two","three"))

```

8, 9, 10. More involved examples here...
